# Telling a Story from data
this is the file for the report exercise 4.4 "telling a story from data"
# 0. General code
```{r}
library(ggplot2)
library(dplyr)
```

# 1. Data import
The data comes from the source:
Chambers, J. M., Cleveland, W. S., Kleiner, B. and Tukey, P. A. (1983) Graphical Methods for Data Analysis. Belmont, CA: Wadsworth.
```{r}
airquality1<-data.frame(airquality)
airquality1<-airquality1|>dplyr::mutate(timeaxis = as.Date(paste(day=airquality1$Day,airquality1$Month,sep="-"),"%d-%m"))
```

# Plots
```{r}
ggplot(data=airquality1, aes(x=timeaxis, y=Temp))+geom_line(size=0.5,color="red")+labs(x="Time in months",y="Temperature in Fahrenheit")+geom_line(aes(y=Ozone),size=0.5,color="blue")

```
in red we have the temperature in Fahrenheit and in blue there is the Ozone in pbb (see help airqualities{datasets})

# Analysis
Next step is to analyze if there is a linear regregression between ozone and temperature
```{r}
lm_ozon_temp<-lm(data=airquality1,formula=airquality1$Ozone~airquality1$Temp)
summary(lm_ozon_temp)
```
The adjusted Rsquared has the value 0.4832 which is quite bad.

# Figures
```{r}
ggplot(data=airquality1,aes(x=Ozone,y=Temp))+
  geom_point(size=0.5)+
  geom_smooth(method="lm", color="purple")
```


```{r}
boxplot(airquality1$Temp~airquality1$Month,data=airquality1,xlab="Months",ylab="Temperature in Fahrenheit")
```

# Description
The graphic shows that there is a korrelation between ozone ande the temperature, but not very strong.
The boxplot shows that in the summer months the temperature are higher than in May.

# Interpretation and discussion
The Ozone can't be only dependent from Temperature, but as shown there is a light correlation. If a causality is given can't be told.