# stepwise
this is the file re_stepwise for the report exercise 8.5
some code chunks/structure are inspired of the book AGDS1 (Pepe Aràn 2023)

## import library
```{r message=FALSE}
library(ggplot2)
library(visdat)
```
## import data
The datasource is: https://raw.githubusercontent.com/geco-bern/agds/main/data/df_for_stepwise_regression.csv
The data got downloaded and stored in the data folder as csv-file.

```{r}
data_stepwise<-read.csv(file="../data/df_for_stepwise_regression_CSV.csv")
```

look at the data with "summary"
```{r}
summary(data_stepwise)
```
so the first two columns aren't numerical, so they aren't usefull for the statisics of numerical data

look at the missing data with "vis_miss"
```{r}
plot1<-vis_miss(data_stepwise)
plot1
```

Here we can see, that there is 5.7% of the data missing.

## Task 1 of the report exercise 8.5

According to the report_exercise task we have to do the steps 1-3 of the chapter 8.2.3.1 of the AGDS1-Book (Pepe Aràn 2023). As step one p has to be 1, step two is to fit the regressions and calculate Rsquared and finally the step 3 is to select the best fitting modell, so the one with the highest Rsquared-value and then compute the AIC.

The variable GPP_NT_VUT_REF is the one that is intersting for our regression model.
```{r}
# for the storage of the calculated values there will be needed some "container" such as for example a vector or a matrix
vector1<-NULL


#for(i in 1:ncol(data_stepwise) ){
 # vector1[i]<-lm(data=data_stepwise,formula=(data_stepwise$GPP_NT_VUT_REF~data_stepwise$TA_F))
#}

#summary(lm(data=data_stepwise,formula=(data_stepwise$GPP_NT_VUT_REF~data_stepwise$TA_F)))

lm(data=data_stepwise,formula=(data_stepwise$GPP_NT_VUT_REF~data_stepwise$TA_F))

```

## Task 2 of the report exercise 8.5

